A family of accessibility measures derived from spatial interaction principles

### 2025-05-26 paper/manuscript.qmd
additional small kappa related typos fixed

$$
B_a = \frac{(P_a\cdot P_T)^N}{D_{aT}^n}
$$ {#eq-reilly}

@@

$$
C^2 = \frac{P_1\cdot P_2}{D_{12}}
$$ {#eq-zipf}

@@

$$
F = G\frac{(\mu_1N_1)(\mu_2N_2)}{d_{12}^2} = G\frac{M_1\cdot M_2}{d_{12}^2} 
$$ {#eq-stewart-force}

@@

\noindent where $D$ is the population density over an infinitesimal area $ds$ and $r$ is the distance to $i$. In @eq-stewart-population-potential-integral, $D\cdot ds$ gives an infinitesimal count of the population, say $dm$, and so, after discretizing space, @eq-stewart-population-potential-integral can be rewritten as:

@@

$$
S_{i} = \sum_j O_j \cdot f(d_{ij})
$$ {#eq-accessibility-general}

@@

The omission of this constant generates a fundamental problem for the measurement unit of accessibility estimates, which undermines the interpretation, communication and comparability of accessibility analysis. Those reading @hansen1959 must recall that @stewartDemographicGravitationEvidence1948 had set the proportionality constant $G$ to 1, with a note that "$G$ \[was\] left for future determination: a suitable choice of other units can reduce it to unity" [p. 34]. In practice, the persistent omission of the constant in accessibility analyses means that $G$ continues to be implicitly set to $1$, even when the fundamental relationship in accessibility is proportionality (e.g., $S_{i} \propto \sum_j g(O_j)f(d_{ij})$) and not equality [for instance, see the formula for accessibility at the top of Figure 1 in @wuUnifyingAccess2020]. The direct consequence is that without a proportionality constant, the units of $S_i$ remain unclear: the unit of "potential of opportunity for interaction" is left free to change as $\beta$ is calibrated. For example, if $c_{ij}$ is distance in meters, it will be number of opportunities per $m^{\beta}$ when $f(c_{ij}) = d^{-\beta}$ but number of opportunities per $e^{-\beta\cdot m}$ when $f(c_{ij}) = e^{-\beta\cdot m}$. This undermines the comparability of accessibility metrics with different decay functions, and renders their results difficult to understand and communicate. The Hansen-style accessibility estimates found in the literature, therefore, are better thought of as ordinal measures of potential that can only be interpreted in terms of higher and lower accessibility, but which has not palpable meaning [@millerAccessibilityMeasurementApplication2018].

@@

The model in @eq-phys-gravity-model posits a quantity $T_{ij}$ that represents a value in a matrix of flows of size $n \times m$, that is, between $i = 1,\cdots, n$ origins and $j = 1,\cdots, m$ destinations. The quantities $W_i^{(1)}$ and $W_j^{(2)}$ are proxies for the masses at $i=1,\cdots,n$ origins and $j=1,\cdots,m$ destinations. The super-indices $(1)$ and $(2)$ are meant to indicate that these masses could be different things, i.e., $W_i^{(1)}$ could be populations, and $W_j^{(2)}$ hectares of park space. Finally, $f(c_{ij})$ is some function of travel cost $c_{ij}$ which reflects travel impedance. In this way, $T_{ij}$ explicitly measures *interaction* in the unit of trips, and the role of $k$ is to ensure that the system-wide sum of $T_{ij}$ represents the total flows in the data. In other words, $k$ is a scale parameter that makes the overall amount of flows identical to the magnitude of the phenomenon being modeled. In other words, it balances the units, in a conceptually similar sense as the gravitational constant in Newton’s Law of Universal Gravitation.

@@

# The product of $A$ and its transpose gives the bibliographical coupling by cited references $B = A\cdot A^T$

@@

Similar to @eq-phys-gravity-model, $W_j^{(2)}$ above is the mass at the destination and the sub-indices are for $i = 1,\cdots, n$ origins and $j = 1,\cdots, m$ destinations.

@@

$$
V^T_{ij} = K^T \cdot W_j^{(2)} \cdot f(c_{ij})
$$ {#eq-total-constrained-access}

@@

$$
V^T_i = \sum_j V^T_{ij} = K^T \sum_j W^{(2)}_jf(c_{ij}) = K^T \cdot V^0_i
$$ {#eq-total-constrained-accessibility}

@@

$$
V^T_i = K^T\sum_j W^{(2)}_jf(c_{ij}) = \sum_j W^{(2)}_j\frac{D\cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
$$

@@

$$
\kappa_{ij}^T = \frac{D\cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
$$

\noindent and therefore $V^T_i$ is now in the units of $W^{(2)}_j$, that is, the mass at the destination ($V^T_i = \sum_j \kappa_{ij}^T W^{(2)}_j$). The role of $\kappa_{ij}^T$ in this reformulation of accessibility is to transform between units and to adjust the number of opportunities accessible from $i$ so they represent a proportion of the total number of opportunities in the region. $\kappa_{ij}^T$ then assigns opportunities in proportion to the impedance between $i$ and $j$. This is why we refer to $\kappa_{ij}^T$ as a proportional allocation factor. On the other hand, the proportionality constant $K^T$ balances the units of $V^0_i$, the Hansen-type accessibility values, and is an alternative expression of the total constrained accessibility measure.

@@

$$
\begin{array}{l}
K^T = \frac{D}{\sum_{i}\sum_{j} W_j^{(2)} f(c_{ij})}\\
K^T = \frac{D}{\frac{W_1^{(2)}}{c_{11}^3}+\frac{W_1^{(2)}}{c_{21}^3} + \frac{W_1^{(2)}}{c_{31}^3} + \cdots + \frac{W_3^{(2)}}{c_{31}^3} + \frac{W_3^{(2)}}{c_{32}^3} + \frac{W_3^{(2)}}{c_{33}^3}
}\\
K^T = \frac{490}{0.6233422} \\
K^T = 786.085
\end{array}
$$

@@

$$
M^T_{ji} = \hat K^T \cdot W_i^{(1)} f(c_{ij}) = \hat K^T \cdot M_{ji}^0
$$ {#eq-total-constrained-market}

@@

$$
\sum_j M^T_{j} = \sum_i\sum_j M^T_{ji} = O
$$ {#eq-total-constraint-market}

@@

$$
\hat K^T=\frac{O}{\sum_i\sum_i M^T_{ji}} = \frac{O}{\sum_i\sum_j W^{(1)}_i f(c_{ij})} 
$$ {#eq-total-population-balancing-factor}

The constrained market potential then takes the following form:

$$
M^T_j = \hat K^T\sum_i W^{(1)}_if(c_{ij}) = \sum_i W^{(2)}_i \frac{O \cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
$$

Where the following $\hat \kappa_{ji}^T$ proportional allocation factor is dimensionless:

$$
\hat \kappa_{ji}^T = \frac{O \cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
$$

Returning back to the numerical example, the proportionality constant $\hat K^T$ is solved for each travel behaviour scenario, and the market potential of each zone $M^T_j$ is expressed as units of population (e.g., the number of people accessible from each origin at that destination) in @tbl-simple-example-total-population-accessibility.

@@

$$
M^S_{j} = \hat \kappa^S_{ji} \sum_i  O_i
$$ {#eq-production-constrained-accessibility-with-proportional-allocation-factor}

>>>>

$$
B_a = \frac{(P_a P_T)^N}{D_{aT}^n}
$$ {#eq-reilly}

@@

$$
C^2 = \frac{P_1 P_2}{D_{12}}
$$ {#eq-zipf}

@@

$$
F = G\frac{(\mu_1N_1)(\mu_2N_2)}{d_{12}^2} = G\frac{M_1 M_2}{d_{12}^2} 
$$ {#eq-stewart-force}

@@

\noindent where $D$ is the population density over an infinitesimal area $ds$ and $r$ is the distance to $i$. In @eq-stewart-population-potential-integral, $D ds$ gives an infinitesimal count of the population, say $dm$, and so, after discretizing space, @eq-stewart-population-potential-integral can be rewritten as:

@@

$$
S_{i} = \sum_j O_j  f(d_{ij})
$$ {#eq-accessibility-general}

@@

The omission of this constant generates a fundamental problem for the measurement unit of accessibility estimates, which undermines the interpretation, communication and comparability of accessibility analysis. Those reading @hansen1959 must recall that @stewartDemographicGravitationEvidence1948 had set the proportionality constant $G$ to 1, with a note that "$G$ \[was\] left for future determination: a suitable choice of other units can reduce it to unity" [p. 34]. In practice, the persistent omission of the constant in accessibility analyses means that $G$ continues to be implicitly set to $1$, even when the fundamental relationship in accessibility is proportionality (e.g., $S_{i} \propto \sum_j g(O_j)f(d_{ij})$) and not equality [for instance, see the formula for accessibility at the top of Figure 1 in @wuUnifyingAccess2020]. The direct consequence is that without a proportionality constant, the units of $S_i$ remain unclear: the unit of "potential of opportunity for interaction" is left free to change as $\beta$ is calibrated. For example, if $c_{ij}$ is distance in meters, it will be number of opportunities per $m^{\beta}$ when $f(c_{ij}) = d^{-\beta}$ but number of opportunities per $e^{-\beta m}$ when $f(c_{ij}) = e^{-\beta m}$. This undermines the comparability of accessibility metrics with different decay functions, and renders their results difficult to understand and communicate. The Hansen-style accessibility estimates found in the literature, therefore, are better thought of as ordinal measures of potential that can only be interpreted in terms of higher and lower accessibility, but which has not palpable meaning [@millerAccessibilityMeasurementApplication2018].

@@

The model in @eq-phys-gravity-model posits a quantity $T_{ij}$ that represents a value in a matrix of flows of size $n \times m$, that is, between $i = 1,s, n$ origins and $j = 1,s, m$ destinations. The quantities $W_i^{(1)}$ and $W_j^{(2)}$ are proxies for the masses at $i=1,s,n$ origins and $j=1,s,m$ destinations. The super-indices $(1)$ and $(2)$ are meant to indicate that these masses could be different things, i.e., $W_i^{(1)}$ could be populations, and $W_j^{(2)}$ hectares of park space. Finally, $f(c_{ij})$ is some function of travel cost $c_{ij}$ which reflects travel impedance. In this way, $T_{ij}$ explicitly measures *interaction* in the unit of trips, and the role of $k$ is to ensure that the system-wide sum of $T_{ij}$ represents the total flows in the data. In other words, $k$ is a scale parameter that makes the overall amount of flows identical to the magnitude of the phenomenon being modeled. In other words, it balances the units, in a conceptually similar sense as the gravitational constant in Newton’s Law of Universal Gravitation.

@@

# The product of $A$ and its transpose gives the bibliographical coupling by cited references $B = A A^T$

@@

Similar to @eq-phys-gravity-model, $W_j^{(2)}$ above is the mass at the destination and the sub-indices are for $i = 1,s, n$ origins and $j = 1,s, m$ destinations.

@@

$$
V^T_{ij} = K^T  W_j^{(2)}  f(c_{ij})
$$ {#eq-total-constrained-access}

@@

$$
V^T_i = \sum_j V^T_{ij} = K^T \sum_j W^{(2)}_jf(c_{ij}) = K^T  V^0_i
$$ {#eq-total-constrained-accessibility}

@@

$$
V^T_i = K^T\sum_j W^{(2)}_jf(c_{ij}) = \sum_j D\frac{W^{(2)}_j f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
$$

@@

$$
\kappa_{ij}^T = \frac{W^{(2)}_j f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
$$

\noindent and therefore $V^T_i$ is now in the units of $D$, that is, the mass at the destination ($V^T_i = \sum_j \kappa_{ij}^T D$). The role of $\kappa_{ij}^T$ in this reformulation of accessibility is to transform between units and to adjust the number of opportunities accessible from $i$ so they represent a proportion of the total number of opportunities in the region $D$. $\kappa_{ij}^T$ then assigns opportunities in proportion to the impedance between $i$ and $j$. This is why we refer to $\kappa_{ij}^T$ as a proportional allocation factor. On the other hand, the proportionality constant $K^T$ balances the units of $V^0_i$, the Hansen-type accessibility values, and is an alternative expression of the total constrained accessibility measure.

@@

$$
\begin{array}{l}
K^T = \frac{D}{\sum_{i}\sum_{j} W_j^{(2)} f(c_{ij})}\\
K^T = \frac{D}{\frac{W_1^{(2)}}{c_{11}^3}+\frac{W_1^{(2)}}{c_{21}^3} + \frac{W_1^{(2)}}{c_{31}^3} + s + \frac{W_3^{(2)}}{c_{31}^3} + \frac{W_3^{(2)}}{c_{32}^3} + \frac{W_3^{(2)}}{c_{33}^3}
}\\
K^T = \frac{490}{0.6233422} \\
K^T = 786.085
\end{array}
$$

@@

$$
M^T_{ji} = \hat K^T  W_i^{(1)} f(c_{ij}) = \hat K^T  M_{ji}^0
$$ {#eq-total-constrained-market}

@@

$$
\sum_j M^T_{j} = \sum_i\sum_j M^T_{ji} = O
$${#eq-total-constraint-market}

@@

$$
\hat K^T=\frac{O}{\sum_i\sum_i M^T_{ji}} = \frac{O}{\sum_i\sum_j W^{(1)}_i f(c_{ij})}
$${#eq-total-population-balancing-factor}

The constrained market potential then takes the following form:
$$
M^T_j = \hat K^T \sum_i W^{(1)}_if(c_{ij}) = O \sum_i \frac{W_i^{(1)} f(c_{ij})}{\sum_i\sum_j W^{(1)}_i f(c_{ij})}
$$

Where the following $\hat\kappa_{ji}^T$ proportional allocation factor is dimensionless, reflecting the proportional allocation of population in the region $O$ (i.e., $M^T_j = \sum_i \hat \kappa_{ji}^T O$) 
$$
\hat \kappa_{ji}^T = \frac{W^{(2)}_i  f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
$$

Returning back to the numerical example, the balancing factor $\hat K^T$ is solved for each travel behaviour scenario, and the market potential of each zone $M^T_j$ is expressed as units of population (e.g., the number of people accessible from each origin at that destination) in @tbl-simple-example-total-population-accessibility.

@@

$$
M^S_{j} = \sum_i \hat \kappa^S_{ji} O_i
$$ {#eq-production-constrained-accessibility-with-proportional-allocation-factor}

### 2025-05-26 paper/manuscript.tex
additional small kappa related typos fixed

\begin{equation}\phantomsection\label{eq-reilly}{
B_a = \frac{(P_a\cdot P_T)^N}{D_{aT}^n}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-zipf}{
C^2 = \frac{P_1\cdot P_2}{D_{12}}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-stewart-force}{
F = G\frac{(\mu_1N_1)(\mu_2N_2)}{d_{12}^2} = G\frac{M_1\cdot M_2}{d_{12}^2} 
}\end{equation}

@@

\noindent where \(D\) is the population density over an infinitesimal
area \(ds\) and \(r\) is the distance to \(i\). In
Equation~\ref{eq-stewart-population-potential-integral}, \(D\cdot ds\)
gives an infinitesimal count of the population, say \(dm\), and so,
after discretizing space,
Equation~\ref{eq-stewart-population-potential-integral} can be rewritten
as:

@@

\begin{equation}\phantomsection\label{eq-accessibility-general}{
S_{i} = \sum_j O_j \cdot f(d_{ij})
}\end{equation}

@@

The omission of this constant generates a fundamental problem for the
measurement unit of accessibility estimates, which undermines the
interpretation, communication and comparability of accessibility
analysis. Those reading Hansen (1959) must recall that Stewart (1948)
had set the proportionality constant \(G\) to 1, with a note that
``\(G\) {[}was{]} left for future determination: a suitable choice of
other units can reduce it to unity'' {[}p.~34{]}. In practice, the
persistent omission of the constant in accessibility analyses means that
\(G\) continues to be implicitly set to \(1\), even when the fundamental
relationship in accessibility is proportionality (e.g.,
\(S_{i} \propto \sum_j g(O_j)f(d_{ij})\)) and not equality (for
instance, see the formula for accessibility at the top of Figure 1 in Wu
and Levinson 2020). The direct consequence is that without a
proportionality constant, the units of \(S_i\) remain unclear: the unit
of ``potential of opportunity for interaction'' is left free to change
as \(\beta\) is calibrated. For example, if \(c_{ij}\) is distance in
meters, it will be number of opportunities per \(m^{\beta}\) when
\(f(c_{ij}) = d^{-\beta}\) but number of opportunities per
\(e^{-\beta\cdot m}\) when \(f(c_{ij}) = e^{-\beta\cdot m}\). This
undermines the comparability of accessibility metrics with different
decay functions, and renders their results difficult to understand and
communicate. The Hansen-style accessibility estimates found in the
literature, therefore, are better thought of as ordinal measures of
potential that can only be interpreted in terms of higher and lower
accessibility, but which has not palpable meaning (E. J. Miller 2018).

@@

The model in Equation~\ref{eq-phys-gravity-model} posits a quantity
\(T_{ij}\) that represents a value in a matrix of flows of size
\(n \times m\), that is, between \(i = 1,\cdots, n\) origins and
\(j = 1,\cdots, m\) destinations. The quantities \(W_i^{(1)}\) and
\(W_j^{(2)}\) are proxies for the masses at \(i=1,\cdots,n\) origins and
\(j=1,\cdots,m\) destinations. The super-indices \((1)\) and \((2)\) are
meant to indicate that these masses could be different things, i.e.,
\(W_i^{(1)}\) could be populations, and \(W_j^{(2)}\) hectares of park
space. Finally, \(f(c_{ij})\) is some function of travel cost \(c_{ij}\)
which reflects travel impedance. In this way, \(T_{ij}\) explicitly
measures \emph{interaction} in the unit of trips, and the role of \(k\)
is to ensure that the system-wide sum of \(T_{ij}\) represents the total
flows in the data. In other words, \(k\) is a scale parameter that makes
the overall amount of flows identical to the magnitude of the phenomenon
being modeled. In other words, it balances the units, in a conceptually
similar sense as the gravitational constant in Newton's Law of Universal
Gravitation.

@@

Similar to Equation~\ref{eq-phys-gravity-model}, \(W_j^{(2)}\) above is
the mass at the destination and the sub-indices are for
\(i = 1,\cdots, n\) origins and \(j = 1,\cdots, m\) destinations.

@@

\begin{equation}\phantomsection\label{eq-total-constrained-access}{
V^T_{ij} = K^T \cdot W_j^{(2)} \cdot f(c_{ij})
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-total-constrained-accessibility}{
V^T_i = \sum_j V^T_{ij} = K^T \sum_j W^{(2)}_jf(c_{ij}) = K^T \cdot V^0_i
}\end{equation}

@@

\[
V^T_i = K^T\sum_j W^{(2)}_jf(c_{ij}) = \sum_j W^{(2)}_j\frac{D\cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

@@

\[
\kappa_{ij}^T = \frac{D\cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

\noindent and therefore \(V^T_i\) is now in the units of \(W^{(2)}_j\),
that is, the mass at the destination
(\(V^T_i = \sum_j \kappa_{ij}^T W^{(2)}_j\)). The role of
\(\kappa_{ij}^T\) in this reformulation of accessibility is to transform
between units and to adjust the number of opportunities accessible from
\(i\) so they represent a proportion of the total number of
opportunities in the region. \(\kappa_{ij}^T\) then assigns
opportunities in proportion to the impedance between \(i\) and \(j\).
This is why we refer to \(\kappa_{ij}^T\) as a proportional allocation
factor. On the other hand, the proportionality constant \(K^T\) balances
the units of \(V^0_i\), the Hansen-type accessibility values, and is an
alternative expression of the total constrained accessibility measure.

@@

\[
\begin{array}{l}
K^T = \frac{D}{\sum_{i}\sum_{j} W_j^{(2)} f(c_{ij})}\\
K^T = \frac{D}{\frac{W_1^{(2)}}{c_{11}^3}+\frac{W_1^{(2)}}{c_{21}^3} + \frac{W_1^{(2)}}{c_{31}^3} + \cdots + \frac{W_3^{(2)}}{c_{31}^3} + \frac{W_3^{(2)}}{c_{32}^3} + \frac{W_3^{(2)}}{c_{33}^3}
}\\
K^T = \frac{490}{0.6233422} \\
K^T = 786.085
\end{array}
\]

@@

\begin{equation}\phantomsection\label{eq-total-constrained-market}{
M^T_{ji} = \hat K^T \cdot W_i^{(1)} f(c_{ij}) = \hat K^T \cdot M_{ji}^0
}\end{equation}

@@

The constrained market potential then takes the following form:

\[
M^T_j = \hat K^T\sum_i W^{(1)}_if(c_{ij}) = \sum_i W^{(2)}_i \frac{O \cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
\]

Where the following \(\hat \kappa_{ji}^T\) proportional allocation
factor is dimensionless:

\[
\hat \kappa_{ji}^T = \frac{O \cdot f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
\]

Returning back to the numerical example, the proportionality constant
\(\hat K^T\) is solved for each travel behaviour scenario, and the
market potential of each zone \(M^T_j\) is expressed as units of
population (e.g., the number of people accessible from each origin at
that destination) in
Table~\ref{tbl-simple-example-total-population-accessibility}.

@@

\begin{equation}\phantomsection\label{eq-production-constrained-accessibility-with-proportional-allocation-factor}{
M^S_{j} = \hat \kappa^S_{ji} \sum_i  O_i
}\end{equation}

>>>>

\begin{equation}\phantomsection\label{eq-reilly}{
B_a = \frac{(P_a P_T)^N}{D_{aT}^n}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-zipf}{
C^2 = \frac{P_1 P_2}{D_{12}}
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-stewart-force}{
F = G\frac{(\mu_1N_1)(\mu_2N_2)}{d_{12}^2} = G\frac{M_1 M_2}{d_{12}^2} 
}\end{equation}

@@

\noindent where \(D\) is the population density over an infinitesimal
area \(ds\) and \(r\) is the distance to \(i\). In
Equation~\ref{eq-stewart-population-potential-integral}, \(D ds\) gives
an infinitesimal count of the population, say \(dm\), and so, after
discretizing space,
Equation~\ref{eq-stewart-population-potential-integral} can be rewritten
as:

@@

\begin{equation}\phantomsection\label{eq-accessibility-general}{
S_{i} = \sum_j O_j  f(d_{ij})
}\end{equation}

@@

The omission of this constant generates a fundamental problem for the
measurement unit of accessibility estimates, which undermines the
interpretation, communication and comparability of accessibility
analysis. Those reading Hansen (1959) must recall that Stewart (1948)
had set the proportionality constant \(G\) to 1, with a note that
``\(G\) {[}was{]} left for future determination: a suitable choice of
other units can reduce it to unity'' {[}p.~34{]}. In practice, the
persistent omission of the constant in accessibility analyses means that
\(G\) continues to be implicitly set to \(1\), even when the fundamental
relationship in accessibility is proportionality (e.g.,
\(S_{i} \propto \sum_j g(O_j)f(d_{ij})\)) and not equality (for
instance, see the formula for accessibility at the top of Figure 1 in Wu
and Levinson 2020). The direct consequence is that without a
proportionality constant, the units of \(S_i\) remain unclear: the unit
of ``potential of opportunity for interaction'' is left free to change
as \(\beta\) is calibrated. For example, if \(c_{ij}\) is distance in
meters, it will be number of opportunities per \(m^{\beta}\) when
\(f(c_{ij}) = d^{-\beta}\) but number of opportunities per
\(e^{-\beta m}\) when \(f(c_{ij}) = e^{-\beta m}\). This undermines the
comparability of accessibility metrics with different decay functions,
and renders their results difficult to understand and communicate. The
Hansen-style accessibility estimates found in the literature, therefore,
are better thought of as ordinal measures of potential that can only be
interpreted in terms of higher and lower accessibility, but which has
not palpable meaning (E. J. Miller 2018).

@@

The model in Equation~\ref{eq-phys-gravity-model} posits a quantity
\(T_{ij}\) that represents a value in a matrix of flows of size
\(n \times m\), that is, between \(i = 1,s, n\) origins and
\(j = 1,s, m\) destinations. The quantities \(W_i^{(1)}\) and
\(W_j^{(2)}\) are proxies for the masses at \(i=1,s,n\) origins and
\(j=1,s,m\) destinations. The super-indices \((1)\) and \((2)\) are
meant to indicate that these masses could be different things, i.e.,
\(W_i^{(1)}\) could be populations, and \(W_j^{(2)}\) hectares of park
space. Finally, \(f(c_{ij})\) is some function of travel cost \(c_{ij}\)
which reflects travel impedance. In this way, \(T_{ij}\) explicitly
measures \emph{interaction} in the unit of trips, and the role of \(k\)
is to ensure that the system-wide sum of \(T_{ij}\) represents the total
flows in the data. In other words, \(k\) is a scale parameter that makes
the overall amount of flows identical to the magnitude of the phenomenon
being modeled. In other words, it balances the units, in a conceptually
similar sense as the gravitational constant in Newton's Law of Universal
Gravitation.

@@

Similar to Equation~\ref{eq-phys-gravity-model}, \(W_j^{(2)}\) above is
the mass at the destination and the sub-indices are for \(i = 1,s, n\)
origins and \(j = 1,s, m\) destinations.

@@

\begin{equation}\phantomsection\label{eq-total-constrained-access}{
V^T_{ij} = K^T  W_j^{(2)}  f(c_{ij})
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-total-constrained-accessibility}{
V^T_i = \sum_j V^T_{ij} = K^T \sum_j W^{(2)}_jf(c_{ij}) = K^T  V^0_i
}\end{equation}

@@

\[
V^T_i = K^T\sum_j W^{(2)}_jf(c_{ij}) = \sum_j D\frac{W^{(2)}_j f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

@@

\[
\kappa_{ij}^T = \frac{W^{(2)}_j f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}
\]

\noindent and therefore \(V^T_i\) is now in the units of \(D\), that is,
the mass at the destination (\(V^T_i = \sum_j \kappa_{ij}^T D\)). The
role of \(\kappa_{ij}^T\) in this reformulation of accessibility is to
transform between units and to adjust the number of opportunities
accessible from \(i\) so they represent a proportion of the total number
of opportunities in the region \(D\). \(\kappa_{ij}^T\) then assigns
opportunities in proportion to the impedance between \(i\) and \(j\).
This is why we refer to \(\kappa_{ij}^T\) as a proportional allocation
factor. On the other hand, the proportionality constant \(K^T\) balances
the units of \(V^0_i\), the Hansen-type accessibility values, and is an
alternative expression of the total constrained accessibility measure.

@@

\[
\begin{array}{l}
K^T = \frac{D}{\sum_{i}\sum_{j} W_j^{(2)} f(c_{ij})}\\
K^T = \frac{D}{\frac{W_1^{(2)}}{c_{11}^3}+\frac{W_1^{(2)}}{c_{21}^3} + \frac{W_1^{(2)}}{c_{31}^3} + s + \frac{W_3^{(2)}}{c_{31}^3} + \frac{W_3^{(2)}}{c_{32}^3} + \frac{W_3^{(2)}}{c_{33}^3}
}\\
K^T = \frac{490}{0.6233422} \\
K^T = 786.085
\end{array}
\]

@@

\begin{equation}\phantomsection\label{eq-total-constrained-market}{
M^T_{ji} = \hat K^T  W_i^{(1)} f(c_{ij}) = \hat K^T  M_{ji}^0
}\end{equation}

@@

The constrained market potential then takes the following form: \[
M^T_j = \hat K^T \sum_i W^{(1)}_if(c_{ij}) = O \sum_i \frac{W_i^{(1)} f(c_{ij})}{\sum_i\sum_j W^{(1)}_i f(c_{ij})}
\]

Where the following \(\hat\kappa_{ji}^T\) proportional allocation factor
is dimensionless, reflecting the proportional allocation of population
in the region \(O\) (i.e., \(M^T_j = \sum_i \hat \kappa_{ji}^T O\)) \[
\hat \kappa_{ji}^T = \frac{W^{(2)}_i  f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}
\]

Returning back to the numerical example, the balancing factor
\(\hat K^T\) is solved for each travel behaviour scenario, and the
market potential of each zone \(M^T_j\) is expressed as units of
population (e.g., the number of people accessible from each origin at
that destination) in
Table~\ref{tbl-simple-example-total-population-accessibility}.

@@

\begin{equation}\phantomsection\label{eq-production-constrained-accessibility-with-proportional-allocation-factor}{
M^S_{j} = \sum_i \hat \kappa^S_{ji} O_i
}\end{equation}

### 2025-05-26 paper/summary-family-access-measures-table.tex
additional small kappa related typos fixed

& Allocates the total marginal as opportunities based on $\kappa_{ij}^T = K^T f(c_{ij})$ and as population based on  $\hat{\kappa}_{ji}^T = \hat K^T f(c_{ij})$

& $V^T_i =  \sum_j \kappa_{ij}^T W^{(2)}_j$;

$M_j^T =  \sum_i \hat{\kappa}_{ji}^T W^{(1)}_i$

@@

& Allocates the single opportunities marginal proportionally based on $\kappa^S_{ij} =  B_j W_i^{(1)} f(c_{ij})$ in the case of $V_i^S$ and the single population marginal proroportionally based on $\hat \kappa^S_{ji} = A_i W_j^{(2)} f(c_{ij})$ in the case of $M_j^S$.

>>>>

& Allocates the total marginal as opportunities $D$ based on $\kappa_{ij}^T = \frac{W^{(2)}_j f(c_{ij})}{\sum_i\sum_j W^{(2)}_jf(c_{ij})}$ and as population $O$ based on $\hat \kappa_{ji}^T = \frac{W^{(2)}_i  f(c_{ij})}{\sum_i\sum_j W^{(2)}_if(c_{ij})}$

& $V^T_i =  \sum_j \kappa_{ij}^T D$;

$M_j^T =  \sum_i \hat{\kappa}_{ji}^T O$

@@

& Allocates the single opportunities marginal $D_j$ proportionally based on $\kappa^S_{ij} =  \frac{W_i^{(1)} f(c_{ij})}{\sum_i W_i^{(1)} f(c_{ij})}$ in the case of $V_i^S$ and the single population marginal $O_i$ proroportionally based on $\hat \kappa^S_{ji} = \frac{W_j^{(2)} f(c_{ij})}{\sum_i W_j^{(2)} f(c_{ij})}$ in the case of $M_j^S$.
